Motivated by the dexterity, terrain scalability and reach availed to humans through the use of legs, and for social and aesthetic reasons, robotics researchers have sought to create humanoid robots. While there are notable examples of significant success in this endeavour, the motion characteristics of these robotic walkers are noticeably different from our own. This is largely due to the restricting of the robotic systems to full actuation, which permits convenient simplifications and robustness. This restriction imposes more than simply visual differences; the conservative motion restricts the speed at which these robots can move and, importantly, their efficiency. Consequently, there exists interest in developing underactuated robotic walkers which utilise the dynamics of the system to generate much of the motion, imposing actuation only where required to enforce a trajectory or respond to disturbance.

However, permitting underactuation causes the path planning and control problems to become significantly more complex. This is mainly due to two contributing factors. Firstly, the dynamics of walking robots are highly nonlinear, particularly about the foot impacts, and so the system model must be very complex or else be a poor approximation. Secondly, the search space to find feasible paths is vast, since in general robotic walkers have many independent degrees of freedom, making traditional methods of planning and control computationally intractable. In addition, control problems of underactuated and fully actuated control are sufficiently different that much of the intuition and methods from the quite well studied field of fully actuated control are not easily applicable to underactuated cases.

Manchester et al \cite{manchester13planning} have proposed a method for path planning for underactuated dynamic walkers using a receding-horizon algorithm which draws from a library of motion primitives defined as virtual holonomic constraints to address the above mentioned challenges. The use of a library of motion primitives drastically reduces the search space and notionally limits the trajectories to feasible motion. The complications regarding the nonlinear impact dynamics are isolated by defining of motion primitives by the duration of one continuous phase – that is, the movement of the robotic walker between one foot impact and the next. Also, the use of virtual holonomic constraints allows for a partial closed-form solution of the dynamical equations to be computed off-line, thereby significantly reducing the required on-line computation in determining dynamic feasibility. This method has been proven to be capable of planning several footsteps ahead of the current position of the robot in a fraction of a second, thus being feasible for real-time control, on a compass-gait walker and more complex 5-link walker.

Power and energy densities pose strict limits on the dimensions and weight of walking robots with present technology. Additionally, ethical considerations in sustainable design guide us to minimise wasteful energy consumption. There are two applicable means by which the energy consumption of an underactuated walking robot may be reduced in the context of motion planning: Minimising the actuator energy output and reducing the computational requirements of the planning algorithm. The virtual constraints approach provides a convincing means of achieving the latter, however the ability of this approach to produce paths close to the true optimal in terms of energy reduction is dependent upon the optimality and coverage of the motion primitives within the library and the manner by which one primitive is preferred over another.

\section{Problem Statement}
In the light of these motivations, the primary aim of this thesis is to produce a method by which optimal primitives may be generated in a procedure which ensures sufficient coverage of the configuration space of the robot. This library is structured to enable rapid traversal in real-time. In addition, an algorithm by which the primitives may be selected rapidly and intelligently is implemented. The combination of these two components enable efficient real-time path planning of underactuated dynamic walkers.

\section{Structure of thesis}
This thesis builds on the work completed by Manchester \& Umenberger \cite{manchester13planning}, which itself arises from a decade of interest in the literature in the method of virtual constraints for control of underactuated walking robots.

\begin{adjustwidth}{1.4cm}{}
\begin{enumerate}[label=\bfseries Chapter \arabic*]
	\setcounter{enumi}{1}
	\item introduces the relevant literature in the planning and control of underactuated robotics, particularly as it applies to walking robots. Particular attention is paid to the method of virtual constraints.
	\item presents the mathematical and technical background required for understanding and applying virtual constraints to underactuated walking robots. In addition, the techniques used for the principled design of virtual constraints as motion primitives are established.
	\item outlines the primary contribution of this thesis work; the method for automatic generation of a library of virtual constraints.
	\item details the planning algorithm, based upon that found in \cite{manchester13planning}, which is used to select a virtual constraint from the library for the purposes of motion planning.
	\item presents the results of testing the validity of the method introduced in Chapter 4, firstly by examining the properties of the virtual constraints themselves and then using a simulator, which is outlined in Appendix \ref{chap:sim}, to assess its efficacy of use in motion planning
	\item discusses the key outcomes raised in Chapter 6 along with the merits of the virtual constraints method
	\item concludes by establishing the contribution of this thesis to the state of the art and outlining the future work required for the application of the methods used in this work in practice to general underactuated walking robots
\end{enumerate}
\end{adjustwidth}